## LINKS
# File to be analyzed: https://github.com/nthun/cocktail-balance/blob/master/cocktail_data.tsv
# Instructions: ...

```{r}
library(readr)      # function(s): read_tsv
library(data.table) # data.table type
library(stringr)    # function(s): str_replace_all
library(tidyr)      # function(s): separate_longer_delim
library(dplyr)      # function(s): arrange, count, filter, n_distinct, rename
```

## 1. Read the data
Read the cocktail dataset from: <https://github.com/nthun/cocktail-balance>

```{r}
my_data = read_tsv('C:/Users/Tatu/Desktop/Deanne - R programming/1_cocktail_bar/cocktail_data.tsv')
my_data = data.table(my_data)
```

## 2. Transform the data table and clean the ingredient variable!
The ingredients are currently in a single cell for each cocktail. It would be better to put them in separate rows. Also, the variable should be cleaned of all quantities (e.g. 1/2 oz, 2 dashes, etc.), and the indicator of alcohol content (e.g. 47.3% abv). You will need to use this cleaned table in later tasks.

```{r}
# some cleaning
my_data$ingredients = as.character(my_data$ingredients) # cast to character (string) type
my_data$ingredients = tolower(my_data$ingredients)      # cast everything to lowercase

# there are some some typos in the columns
my_data$ingredients = str_replace_all(my_data$ingredients, "<br>", "<br/>")            # lines 9 and 20
my_data$ingredients = str_replace_all(my_data$ingredients, "cr\u008fme", "crème")      # line 17
my_data$ingredients = str_replace_all(my_data$ingredients, "<b4/>", "<br/>")           # line 25
my_data$ingredients = str_replace_all(my_data$ingredients, "marschino", "maraschino")  # line 25
my_data$ingredients = str_replace_all(my_data$ingredients, "simple<", "simple syrup<") # line 54

# TIDYR CHEATSHEET
# https://github.com/rstudio/cheatsheets/blob/main/tidyr.pdf
my_data = separate_longer_delim(my_data, ingredients, "<br/>") # create new rows based on line breaks indicated by "<br/>" inside the ingredients column
my_data = separate_longer_delim(my_data, ingredients, " or ")  # to handle line 53

# regex black magic which I love :)
my_data$ingredients = str_replace_all(my_data$ingredients, "^ +", "") # remove leading whitespaces
my_data$ingredients = str_replace_all(my_data$ingredients, " +$", "") # remove trailing whitespaces

my_data$ingredients = str_replace_all(my_data$ingredients, " \\(\\d+% abv\\)", "")        # remove alcohol "(XX% abv)"
my_data$ingredients = str_replace_all(my_data$ingredients, " \\(\\d+\\.\\d+% abv\\)", "") # remove alcohol " (XX.XX% abv)"

my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+ dash(es)? ", "") # remove "XX dash(es) "

my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+ drop(s)? ", "") # remove "XX drop(s) "

my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+ bsp ", "") # remove "XX bsp "

my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+\\.\\d+ oz ", "")    # remove "XX.XX oz "
my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+ \\d+/\\d+ oz ", "") # remove "XX XX/XX oz "
my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+/\\d+ oz ", "")      # remove "XX/XX oz "
my_data$ingredients = str_replace_all(my_data$ingredients, "\\d+ oz ", "")           # remove "XX oz "

my_data$ingredients = str_replace_all(my_data$ingredients, " +$", "") # remove trailing whitespaces again, because somewhere I messed up... -,-
```

## 3. All ingredients in alphabetical order
Before opening the bar, you need to find a reliable supplier that has all the ingredients. You need to send a list of all possible ingredients you will need. They don't need the quantities (i.e. how many of these are needed), just the names of the ingredients.

```{r}
sort(unique(my_data$ingredients))
```

## 4. Number of unique ingredients
How many different ingredients you will need?

```{r}
length(unique(my_data$ingredients))
```

## 5. What are the top 10 ingredients?
What are the 10 most frequently used ingredients? If there are ties, you can list more than 10.

```{r}
top_10_ingredients = my_data %>%
  count(ingredients) %>% # count the number of rows per ingredients => results in a new column called "n"
  arrange(desc(n))       # sort column "n" in descending order, so the most used ingredients appear on the top

head(top_10_ingredients$ingredients, 10)
```

## 6. Which cocktail(s) has/have the most ingredients?
Count the number of ingredients and filter all the cocktails that has that many.

```{r}
cocktail_with_most_ingredients = my_data %>%
  count(name) %>%        # count the number of rows per cocktail names => results in a new column called "n"
  filter(n == max(n)) %>% # filter the table to only those rows which have the maximal value from "n"
  select(name)

cocktail_with_most_ingredients
```

## 7. How many ingredients appear in only one cocktail (rare ingredient)?
Count all cocktails with only one ingredient

```{r}
rare_ingredients = my_data %>%
  count(ingredients) %>%  # count the number of cocktails per ingredient => results in a new column called "n"
  filter(n == 1)          # filter the table to only those ingredients which have only one cocktail count

rare_ingredients = rare_ingredients$ingredients  # drop count column, keep ingredients only
rare_ingredients
```

## 8. Which cocktail has an ingredient that is only used in one cocktail?

```{r}
cocktails_with_rare_ingredients = my_data %>%
  filter(ingredients %in% rare_ingredients) 

cocktails_with_rare_ingredients = cocktails_with_rare_ingredients$name
cocktails_with_rare_ingredients
```

## 9. What are the cocktails without rare ingredients?

```{r}
cocktails_with_common_ingredients = my_data %>%
  filter(name %notin% cocktails_with_rare_ingredients) %>%
  distinct(name)

cocktails_with_common_ingredients = cocktails_with_common_ingredients$name
cocktails_with_common_ingredients
```

## 10. Create a cheat sheet for the bartender!

Create a matrix that shows all cocktail names as rows and all ingredients as columns. When a cocktail requires an ingredient, there should be an "X" in the cell, otherwise, the cell should remain empty. Example:

```{r echo=FALSE}
tribble(~name, ~Lime, ~`White rum`, ~`Cane sugar`, ~`Coca-cola`, ~Mint, ~Soda,
        "Caipirissima", "X", "X", "X", "", "", "",
        "Cuba Libre", "X", "X", "", "X","","",
        "Mojito", "X", "X", "", "", "X", "X") %>% 
    relocate(name, sort(names(.)))
```

```{r}
my_data$ingredients = str_replace_all(my_data$ingredients, " ", "_")
my_data$mark = "X"

cheatsheet = my_data %>%
  select(name, ingredients, mark) %>%
  pivot_wider(
    names_from = ingredients,
    values_from = mark
  ) %>%
  relocate(name, sort(names(.)))

cheatsheet
```